
\thispagestyle{empty}

\begin{center}
    \LARGE \textbf{Acknowledgements}
\end{center}

\vspace{0.5cm}

\noindent
The success and outcome of this thesis required a lot of guidance and assistance from many people, and I am extremely lucky to come across them during the thesis. I would like to take this opportunity to thank them for their guidance and support in everything I have accomplished. I owe my deep gratitude to my supervisor and thesis guide \textbf{Prof. Debotosh Bhattacharjee}, who was always there during my whole thesis work and guided me all along till the completion of my thesis work by providing all the necessary information that would be beneficial for my research and experiments. I am extremely thankful to him for his assistance and dedicated involvement in every step of the process. His constant suggestions have been an everlasting source of motivation and inspiration to me. It has been an honour to work under a person with wisdom yet a kind, simplistic attitude.

\vspace{0.5cm}
\noindent
I would also like to thank \textbf{Prof. Nirmalya Chowdhury}, Head of the department, Department of Computer Science and Engineering at Jadavpur University, for extending her help to complete this thesis.

\vspace{0.5cm}
\noindent
I had the pleasure of working in the CMATER lab of Jadavpur University. I acknowledge the generous collaboration and assistance of every member of the CMATER lab. It has been a wonderful experience working with every member of this lab. I learnt many things regarding research works and was inspired to do better in my thesis work from them.

\vspace{0.5cm}
\noindent
I especially like to thank my classmate and one of the members of CMATER lab, \textbf{Akash Nayak}, who helped me in many steps of my thesis work.

\vspace{0.5cm}
\noindent
At the end, I like to acknowledge both of my parents for their constant support during this thesis work.

\vspace{1.5cm}
\noindent
\textbf{Nayan Mahata} \\
Examination Roll No.: M4CSE250xx \\
Registration No.: 1661828 of 2023 - 2024 \\
Department of Computer Science and Engineering \\
Jadavpur University

